---
title: "Covid91 vaccine study Final2024"
author: "Mikaela Elias"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
## Methods

Question: How effective is the new preventative vaccine on male individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in males.         
$H_A$ This vaccine is effective in preventing covid in males. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         



```{r}
Male <- FinalDat[FinalDat$sex=='M',]
```   
     
## Graphical Descriptive Results    
```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```
         
The Graphical descriptive results shows that males who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19. 
         
## Numerical Descriptive Results
         
```{r}
table1 <- xtabs(~infected +treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
         
The numerical results shows that the males who received the vaccine, 26.20 have contracted covid, while 49.9 percent have not.
The numerical results also shows that the individuals who received the placebo,73.80 percent have contracted covid, while 50.01 have not. This means that the preventative vaccine is effective in preventing covid 19 in males.
         
## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```         
        
The inferential results show that the Chi squared test shows the p-value of 0.00000000000000022.         
The fishers test shows that the p value is 0.00000000000000022. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated men are 2.3 more times likely to not contract covid than unvaccinated men 

## Conclusion        
Question: How effective is the new preventative vaccine on male individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in males.         
$H_A$ This vaccine is effective in preventing covid in males. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT 

We will reject the null hypothesis because...          
The Graphical descriptive results shows that males who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19.          
The numerical results shows that the males who received the vaccine, 26.20 have contracted covid, while 49.9 percent have not.
The numerical results also shows that the individuals who received the placebo,73.80 percent have contracted covid, while 50.01 have not. This means that the preventative vaccine is effective in preventing covid 19 in males.        
The inferential results show that the Chi squared test shows the p-value of 0.00000000000000022.         
The fishers test shows that the p value is 0.00000000000000022. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.   
The fishers exact test also shows that vaccinated men are 2.3 more times likely to not contract covid than unvaccinated men        

# Females     
## Methods         
        
Question: How effective is the new preventative vaccine on female individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in females.         
$H_A$ This vaccine is effective in preventing covid in females. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         



```{r}
Female <- FinalDat[FinalDat$sex=='F',]
```   
      

## Graphical Descriptive Results    
```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```
          
The Graphical descriptive results shows that females who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19. 
         
## Numerical Descriptive Results
         
```{r}
table1 <- xtabs(~infected +treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```
        
The numerical results shows that the Females who received the vaccine, 35.07 percent have contracted covid, while 50.06 percent have not.
The numerical results also shows that the individuals who received the placebo,64.93 percent have contracted covid, while 49.94 have not. This means that the preventative vaccine is effective in preventing covid 19 in Females.
         
## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```         
The inferential results show that the Chi squared test shows the p-value of 0.00000001914.         
The fishers test shows that the p value is 0.00000001246. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated females are 1.8 more times likely to not contract covid than unvaccinated females.

## Conclusion        
        
Question: How effective is the new preventative vaccine on female individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in females.         
$H_A$ This vaccine is effective in preventing covid in females. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         
     
We will reject the null hypothesis because...             
The Graphical descriptive results shows that females who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19.           
The numerical results shows that the Females who received the vaccine, 35.07 percent have contracted covid, while 50.06 percent have not.
The numerical results also shows that the individuals who received the placebo,64.93 percent have contracted covid, while 49.94 have not. This means that the preventative vaccine is effective in preventing covid 19 in Females.
The inferential results show that the Chi squared test shows the p-value of 0.00000001914.         
The fishers test shows that the p value is 0.00000001246. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated females are 1.8 more times likely to not contract covid than unvaccinated females.   
          
# LGBTQ
## Methods         
        
Question: How effective is the new preventative vaccine on LGBTQ individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in LGBTQ individuals.         
$H_A$ This vaccine is effective in preventing covid in LGBTQ individuals. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         
         

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ=='gay',]
```   

## Graphical Descriptive Results    
```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```
        
The Graphical descriptive results shows that LGBTQ individuals who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19.
         
## Numerical Descriptive Results
         
```{r}
table1 <- xtabs(~infected +treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```        
      
The numerical results shows that the LGBTQ individuals who received the vaccine, 20.00 percent have contracted covid, while 51.67 percent have not.
The numerical results also shows that the individuals who received the placebo,80.00 percent have contracted covid, while 48.33 have not. This means that the preventative vaccine is effective in preventing covid 19 in LGBTQ individuals. 
        
## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```       
      
The inferential results show that the Chi squared test shows the p-value of 0.000002266.         
The fishers test shows that the p value is 0.0000007269. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated LGBTQ individuals are 4.2 more times likely to not contract covid than unvaccinated LGBTQ individuals.
      
## Conclusion 
        
Question: How effective is the new preventative vaccine on LGBTQ individuals?            
          
$H_O$ This vaccine is not effective in preventing covid in LGBTQ individuals.         
$H_A$ This vaccine is effective in preventing covid in LGBTQ individuals. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         
         
We will reject the null hypothesis because...       
The Graphical descriptive results shows that LGBTQ individuals who have taken the preventative vaccine are less likely to contract covid 19, while those who have taken the placebo are more likely to contract covid 19.       
The numerical results shows that the LGBTQ individuals who received the vaccine, 20.00 percent have contracted covid, while 51.67 percent have not.
The numerical results also shows that the individuals who received the placebo,80.00 percent have contracted covid, while 48.33 have not. This means that the preventative vaccine is effective in preventing covid 19 in LGBTQ individuals.        
The inferential results show that the Chi squared test shows the p-value of 0.000002266.         
The fishers test shows that the p value is 0.0000007269. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated LGBTQ individuals are 4.2 more times likely to not contract covid than unvaccinated LGBTQ individuals.     
         
# Druggies      
## Methods         
        
Question: How effective is the new preventative vaccine in drug users?            
          
$H_O$ This vaccine is not effective in preventing covid in drug users.         
$H_A$ This vaccine is effective in preventing covid in drug users. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         
         

```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=='yes',]
```   

## Graphical Descriptive Results    
```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```
        
The Graphical descriptive results shows that drug users who have taken the preventative vaccine are more likely to contract covid 19, while those who have taken the placebo are less likely to contract covid 19.
         
## Numerical Descriptive Results
         
```{r}
table1 <- xtabs(~infected +treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```               
       
The numerical results shows that drug users who received the vaccine, 91.67 percent have contracted covid, while 49.00 percent have not.
The numerical results also shows that the individuals who received the placebo,8.33 percent have contracted covid, while 51.00 have not. This means that the preventative vaccine is not effective in preventing covid 19 in Drug Users. 
        
## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```    

The inferential results show that the Chi squared test shows the p-value of 0.0000000001414.         
The fishers test shows that the p value is 0.000000000003261. Because both of the p-values are under the acceptable rate of 0.5 we will reject the null hypothesis.       
The fishers exact test also shows that vaccinated LGBTQ individuals are 0.08 more times likely to not contract covid than unvaccinated drug users
       
## Conclusion 
        
Question: How effective is the new preventative vaccine in drug users?            
          
$H_O$ This vaccine is not effective in preventing covid in drug users.         
$H_A$ This vaccine is effective in preventing covid in drug users. 
         
infected~sex       
categorical`categorical          
So we will use the helper file CAT~CAT         
         
We will    the null hypotheis because...       
